---
title: "random_forest_"
author: Victor Chuol
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
# Potential packages to be used
library(tidyverse)
#library(arules)
library(caret)
#library(markdown)
library(plyr)
library(dplyr)
library(rattle)
#library(mdsr)
library(mosaic)
#library(mclust)
#library(animation)
#library(flexclust)
library(shuffle)
library(rpart)
library(C50)
library(stats)
```

```{r}
rockData<- read.csv(file = '~/Documents/School Work/Machine Learning/projectdata/0.csv', stringsAsFactors = TRUE)

scissorsData<- read.csv(file = '~/Documents/School Work/Machine Learning/projectdata/1.csv', stringsAsFactors = TRUE)

paperData<- read.csv(file = '~/Documents/School Work/Machine Learning/projectdata/2.csv', stringsAsFactors = TRUE)

okData<- read.csv(file = '~/Documents/School Work/Machine Learning/projectdata/3.csv', stringsAsFactors = TRUE)

```


```{r}
# Rename the columns to show the corresponding sensor and time stamp
names <- c("Sens1_0ms", "Sens2_0ms", "Sens3_0ms", "Sens4_0ms", "Sens5_0ms", "Sens6_0ms", "Sens7_0ms", "Sens8_0ms",
           "Sens1_5ms", "Sens2_5ms", "Sens3_5ms", "Sens4_5ms", "Sens5_5ms", "Sens6_5ms", "Sens7_5ms", "Sens8_5ms",
           "Sens1_10ms","Sens2_10ms","Sens3_10ms","Sens4_10ms","Sens5_10ms","Sens6_10ms","Sens7_10ms","Sens8_10ms",
           "Sens1_15ms","Sens2_15ms","Sens3_15ms","Sens4_15ms","Sens5_15ms","Sens6_15ms","Sens7_15ms","Sens8_15ms",
           "Sens1_20ms","Sens2_20ms","Sens3_20ms","Sens4_20ms","Sens5_20ms","Sens6_20ms","Sens7_20ms","Sens8_20ms",
           "Sens1_25ms","Sens2_25ms","Sens3_25ms","Sens4_25ms","Sens5_25ms","Sens6_25ms","Sens7_25ms","Sens8_25ms",
           "Sens1_30ms","Sens2_30ms","Sens3_30ms","Sens4_30ms","Sens5_30ms","Sens6_30ms","Sens7_30ms","Sens8_30ms",
           "Sens1_35ms","Sens2_35ms","Sens3_35ms","Sens4_35ms","Sens5_35ms","Sens6_35ms","Sens7_35ms","Sens8_35ms",
           "type")

# Apply the new column names
names(rockData) <- names
names(scissorsData) <- names
names(paperData) <- names
names(okData) <- names
```

```{r}
# Renumber the types so rock=1, scissors=2, paper=3, and ok=4
rockData$type[rockData$type == "0"] <- "1"
scissorsData$type[scissorsData$type == "1"] <- "2"
paperData$type[paperData$type == "2"] <- "3"
okData$type[okData$type == "3"] <- "4"
```

```{r}
gestureData <- rbind(rockData,scissorsData,paperData,okData)
#proccesing error handle
shuffledData <- shuffle(gestureData)

# Create the testing and training set
dt = sort(sample(nrow(shuffledData), nrow(shuffledData)*.90))
gesture_train <- shuffledData[dt,]
gesture_test <- shuffledData[-dt,]

#contains just orig.id and type
#gesture_train_no_type <- select(gesture_train,-type,-orig.id)
#head(gesture_train_no_type,3)

gesture_test_no_type <- select(gesture_train,-type,-orig.id)

```

```{r TOTAL DATA CREATION}
#consolidate data into total
tot_gestureData <- gestureData %>% 
  mutate(
    TotSens1 = (Sens1_0ms+Sens1_5ms+Sens1_10ms+Sens1_15ms+Sens1_20ms+Sens1_25ms+Sens1_30ms+Sens1_35ms),
    TotSens2 = (Sens2_0ms+Sens2_5ms+Sens2_10ms+Sens2_15ms+Sens2_20ms+Sens2_25ms+Sens2_30ms+Sens2_35ms),
    TotSens3 = (Sens3_0ms+Sens3_5ms+Sens3_10ms+Sens3_15ms+Sens3_20ms+Sens3_25ms+Sens3_30ms+Sens3_35ms),
    TotSens4 = (Sens4_0ms+Sens4_5ms+Sens4_10ms+Sens4_15ms+Sens4_20ms+Sens4_25ms+Sens4_30ms+Sens4_35ms),
    TotSens5 = (Sens5_0ms+Sens5_5ms+Sens5_10ms+Sens5_15ms+Sens5_20ms+Sens5_25ms+Sens5_30ms+Sens5_35ms),
    TotSens6 = (Sens6_0ms+Sens6_5ms+Sens6_10ms+Sens6_15ms+Sens6_20ms+Sens6_25ms+Sens6_30ms+Sens6_35ms),
    TotSens7 = (Sens7_0ms+Sens7_5ms+Sens7_10ms+Sens7_15ms+Sens7_20ms+Sens7_25ms+Sens7_30ms+Sens7_35ms),
    TotSens8 = (Sens8_0ms+Sens8_5ms+Sens8_10ms+Sens8_15ms+Sens8_20ms+Sens8_25ms+Sens8_30ms+Sens8_35ms),
  ) %>% 
  select(TotSens1,TotSens2,TotSens3,TotSens4,TotSens5,TotSens6,TotSens7,TotSens8,type)

```

```{r TOTAL TRAIN/TEST CREATION}
#create tot data / shuffle test 
shuffle_tot_data <- shuffle(tot_gestureData)


dt2 = sort(sample(nrow(shuffle_tot_data), nrow(shuffle_tot_data)*.70))
gesture_tot_train <- shuffle_tot_data[dt2,]
gesture_tot_test <- shuffle_tot_data[-dt2,]

gesture_tot_test <- select(gesture_tot_test,-orig.id)
gesture_tot_train <- select(gesture_tot_train,-orig.id)
gesture_tot_train_notype <- select(gesture_tot_train,-type)


gesture_tot_test_notype <- select(gesture_tot_test,-type)

```

```{r COMBINED DATA CREATION}
#Split each individual row into 8 new rows based on time. Since each row contains 64 readings from 8 sensors over 40ms, split them on their 5ms intervals so a new row is made for 0ms, 5ms, etc. Do this for all rows in the data set

data0ms <- gestureData %>% 
  select(Sens1_0ms, Sens2_0ms, Sens3_0ms, Sens4_0ms, Sens5_0ms, Sens6_0ms, Sens7_0ms, Sens8_0ms, type) %>%
  rename(replace = c("Sens1_0ms" = "Sens1", "Sens2_0ms" = "Sens2","Sens3_0ms" = "Sens3","Sens4_0ms" = "Sens4","Sens5_0ms" = "Sens5","Sens6_0ms" = "Sens6","Sens7_0ms" = "Sens7", "Sens8_0ms" = "Sens8" ))

data5ms <- gestureData %>% 
  select(Sens1_5ms, Sens2_5ms, Sens3_5ms, Sens4_5ms, Sens5_5ms, Sens6_5ms, Sens7_5ms, Sens8_5ms, type) %>%
  rename(replace = c("Sens1_5ms" = "Sens1", "Sens2_5ms" = "Sens2","Sens3_5ms" = "Sens3","Sens4_5ms" = "Sens4","Sens5_5ms" = "Sens5","Sens6_5ms" = "Sens6","Sens7_5ms" = "Sens7", "Sens8_5ms" = "Sens8" ))

data10ms <- gestureData %>%
  select(Sens1_10ms, Sens2_10ms, Sens3_10ms, Sens4_10ms, Sens5_10ms, Sens6_10ms, Sens7_10ms, Sens8_10ms, type) %>% 
  rename(replace = c("Sens1_10ms" = "Sens1", "Sens2_10ms" = "Sens2","Sens3_10ms" = "Sens3","Sens4_10ms" = "Sens4","Sens5_10ms" = "Sens5","Sens6_10ms" = "Sens6","Sens7_10ms" = "Sens7", "Sens8_10ms" = "Sens8" ))

data15ms <- gestureData %>% 
  select(Sens1_15ms, Sens2_15ms, Sens3_15ms, Sens4_15ms, Sens5_15ms, Sens6_15ms, Sens7_15ms, Sens8_15ms, type) %>% 
  rename(replace = c("Sens1_15ms" = "Sens1", "Sens2_15ms" = "Sens2","Sens3_15ms" = "Sens3","Sens4_15ms" = "Sens4","Sens5_15ms" = "Sens5","Sens6_15ms" = "Sens6","Sens7_15ms" = "Sens7", "Sens8_15ms" = "Sens8" ))

data20ms <- gestureData %>%
  select(Sens1_20ms, Sens2_20ms, Sens3_20ms, Sens4_20ms, Sens5_20ms, Sens6_20ms, Sens7_20ms, Sens8_20ms, type) %>%
  rename(replace = c("Sens1_20ms" = "Sens1", "Sens2_20ms" = "Sens2","Sens3_20ms" = "Sens3","Sens4_20ms" = "Sens4","Sens5_20ms" = "Sens5","Sens6_20ms" = "Sens6","Sens7_20ms" = "Sens7", "Sens8_20ms" = "Sens8" ))

data25ms <- gestureData %>% 
  select(Sens1_25ms, Sens2_25ms, Sens3_25ms, Sens4_25ms, Sens5_25ms, Sens6_25ms, Sens7_25ms, Sens8_25ms, type) %>% 
  rename(replace = c("Sens1_25ms" = "Sens1", "Sens2_25ms" = "Sens2","Sens3_25ms" = "Sens3","Sens4_25ms" = "Sens4","Sens5_25ms" = "Sens5","Sens6_25ms" = "Sens6","Sens7_25ms" = "Sens7", "Sens8_25ms" = "Sens8" ))

data30ms <- gestureData %>%
  select(Sens1_30ms, Sens2_30ms, Sens3_30ms, Sens4_30ms, Sens5_30ms, Sens6_30ms, Sens7_30ms, Sens8_30ms, type) %>%
  rename(replace = c("Sens1_30ms" = "Sens1", "Sens2_30ms" = "Sens2","Sens3_30ms" = "Sens3","Sens4_30ms" = "Sens4","Sens5_30ms" = "Sens5","Sens6_30ms" = "Sens6","Sens7_30ms" = "Sens7", "Sens8_30ms" = "Sens8" ))

data35ms <- gestureData %>%
  select(Sens1_35ms, Sens2_35ms, Sens3_35ms, Sens4_35ms, Sens5_35ms, Sens6_35ms, Sens7_35ms, Sens8_35ms, type) %>% 
  rename(replace = c("Sens1_35ms" = "Sens1", "Sens2_35ms" = "Sens2","Sens3_35ms" = "Sens3","Sens4_35ms" = "Sens4","Sens5_35ms" = "Sens5","Sens6_35ms" = "Sens6","Sens7_35ms" = "Sens7", "Sens8_35ms" = "Sens8" ))

gestureDataCombined <- rbind(data0ms, data5ms, data10ms, data15ms, data20ms, data25ms, data30ms, data35ms)

```

```{r COMBINED TRAIN/TEST CREATION}
## Creating training and testing sets out of spread gesture set```{r}
# Set the seed so this can be reproducible
set.seed(48)

# Shuffle the gesture data before taking training and testing subsets
shuffledData2 <- shuffle(gestureDataCombined)

# Create the testing and training set
dt2 = sort(sample(nrow(shuffledData2), nrow(shuffledData2)*.90))
gest_combine_train <- shuffledData2[dt2,]
gest_combine_test <- shuffledData2[-dt2,]

# Remove the columns denoting the known classification and prior row ID in the training set before creating a model
gest_combine_train_no_type <- gest_combine_train %>% select(-type, -orig.id)

# Remove the original id from the training set that still has type in it
gest_combine_train <- gest_combine_train %>% select(-orig.id)

# Remove the type from the test set
gest_combine_test_no_type <- gest_combine_test %>% select(-type)
```

## Creating a new gesture set containing the averages among each of the 8 sensors so there are now only 8 attributes for each of the sensors
```{r AVG DATA CREATION}
Avg_gestureData <- gestureData %>% 
  mutate(
    AvgSens1 = (Sens1_0ms+Sens1_5ms+Sens1_10ms+Sens1_15ms+Sens1_20ms+Sens1_25ms+Sens1_30ms+Sens1_35ms)/8,
    AvgSens2 = (Sens2_0ms+Sens2_5ms+Sens2_10ms+Sens2_15ms+Sens2_20ms+Sens2_25ms+Sens2_30ms+Sens2_35ms)/8,
    AvgSens3 = (Sens3_0ms+Sens3_5ms+Sens3_10ms+Sens3_15ms+Sens3_20ms+Sens3_25ms+Sens3_30ms+Sens3_35ms)/8,
    AvgSens4 = (Sens4_0ms+Sens4_5ms+Sens4_10ms+Sens4_15ms+Sens4_20ms+Sens4_25ms+Sens4_30ms+Sens4_35ms)/8,
    AvgSens5 = (Sens5_0ms+Sens5_5ms+Sens5_10ms+Sens5_15ms+Sens5_20ms+Sens5_25ms+Sens5_30ms+Sens5_35ms)/8,
    AvgSens6 = (Sens6_0ms+Sens6_5ms+Sens6_10ms+Sens6_15ms+Sens6_20ms+Sens6_25ms+Sens6_30ms+Sens6_35ms)/8,
    AvgSens7 = (Sens7_0ms+Sens7_5ms+Sens7_10ms+Sens7_15ms+Sens7_20ms+Sens7_25ms+Sens7_30ms+Sens7_35ms)/8,
    AvgSens8 = (Sens8_0ms+Sens8_5ms+Sens8_10ms+Sens8_15ms+Sens8_20ms+Sens8_25ms+Sens8_30ms+Sens8_35ms)/8,
  ) %>% 
  select(AvgSens1,AvgSens2,AvgSens3,AvgSens4,AvgSens5,AvgSens6,AvgSens7,AvgSens8,type)
```

## Creating training and testing sets out of averaged gesture set
```{r AVG DATA TRAIN/TEST CREATION}
set.seed(48)

# Shuffle the gesture data before taking training and testing subsets
Avg_shuffledData <- shuffle(Avg_gestureData)

# Create the testing and training set
#dt2 = sort(sample(nrow(Avg_shuffledData), nrow(Avg_shuffledData)*.90))
dt1 = sort(sample(nrow(Avg_shuffledData), nrow(Avg_shuffledData)*.70))

avg_gesture_train <- Avg_shuffledData[dt1,]
avg_gesture_test <- Avg_shuffledData[-dt1,]

# Remove the columns denoting the know classification and prior row ID before creating a model
avg_gesture_train_no_type <- avg_gesture_train %>% 
  select(-type, -orig.id)
avg_gesture_test_no_type <- avg_gesture_test %>% 
  select(-type, -orig.id)
```


```{r GET_FIT FUNCTION}
#RETURNS RPART OR C.50 FIT
getfit <- function(new_formula,data,type){
  #Rpart Tree
  if(type==0){
    formula.fit <- train(new_formula, 
                         data = data,
                         method = "rpart2"
    )
  }
  
  #C5.0 Tree
  if(type == 1){
    formula.fit<-C5.0(new_formula,data,type="class")
  }
  return(formula.fit)  
}

```

```{r GET_FORMULA FUNCTION}
#RETURNS A FORMULA W/ RANDOM SELECTION OF FEATURES
getformula <- function(n.feat,features,targetft,random){
  
  #set number of features to use 
  feature_m <- 6
  
  #random feature selection 
  
  if(random==0){
      #random reduce # selection
      reduce <- runif(1,0.45,1)
      #random # of features implementation 
      sample_features <- sample(features, size = (floor(n.feat) * reduce),replace =F)
  }
  #static # of features
  if(random==1){
      sample_features <- sample(features, size = (feature_m),replace =F)
  }

  #format selected features 
  new_formula <- as.formula(paste0(targetft, " ~ ", paste0(sample_features, collapse =  " + ")))
  
  #tracks completion of function
  print(new_formula)
  
  return (new_formula)
}

```

```{r BIND PREDICTIONS TABLE}
#BINDS MULTIPLE DECISION TREES
bind_predictions <- function(num_predict,test,trees){
  
  list <- vector(mode = "list", length = num_predict)
  
    for (i in 1:num_predict){
      predict.tree <- predict(trees[[i]],test)
      list[[i]] <- predict.tree
    }

  return(list)
}
```

```{r GET N TREES FUNCTION}
get_n_trees <- function(ntrees,train_set,nsamples,features,targetft,type,random){
  
  list <- vector(mode = "list", length = ntrees)
  
  for(i in 1:ntrees){
    train <-sample(train_set,nsamples)
      
      if(missing(random)){
          random <- 1
      }
    test_formula <- getformula(8,features = features,targetft=targetft,random)
      
    new_fit <- getfit(test_formula,train,type)
    list[[i]] <- new_fit
    
  }
  
  return(list)
}
```



```{r GETMEAN FUNCTION}
get_mean<-function(prediction_list){
  col=1
  dataFrame <- c()
  means <- c()

    for(col in 1:length(prediction_list[[1]])){
    
        for(row in 1:npredictions){
            dataFrame <- c(dataFrame,prediction_list[[row]][col])
            }
    mean_ <- mean(dataFrame)
    means <- c(means,rndUp(mean_))
    dataFrame <- c()
    
  }
  means <- as.data.frame(means)
  return(means)
}


```

```{r ROUND UP HELPER FUNCTION}
#rounding up from .5
rndUp <- function(x) trunc(x+sign(x)*0.5)
```


```{r CREATE FORMULA}
formula_combine <- type ~ Sens1+Sens2+Sens3+Sens4+
  Sens5 + Sens6 + Sens7 + Sens8

formula_total <- type ~ TotSens1+TotSens2+TotSens3+TotSens4+
  TotSens5 + TotSens6 + TotSens7 + TotSens8

formula_avg <- type ~ AvgSens1+AvgSens2+AvgSens3+AvgSens4+AvgSens5+
  AvgSens6+AvgSens7+AvgSens8

```

```{r CREATE FEATURES FOR TABLE}
features_t <- all.vars(formula_total)[-1]
targetft_t <- all.vars(formula_total)[1] 

features_a <- all.vars(formula_avg)[-1]
targetft_a <- all.vars(formula_avg)[1] 


features <- all.vars(formula_combine)[-1]
targetft <- all.vars(formula_combine)[1] 

```

```{r SET NPREDICTIONS}
#type = 1 --> C5.0 : type = 0 --> Rpart 
npredictions = 3
nsamples = 2000
```



```{r CREATE SPREAD N TREES}
#SPREAD DATA

#required for C5.0
gest_combine_train$type <- as.factor(gest_combine_train$type)


test_trees_rpart<- get_n_trees(npredictions,gest_combine_train,nsamples,features,targetft,type=0,random=1)
test_trees_c50<- get_n_trees(npredictions,gest_combine_train,nsamples,features,targetft,type=1)

#get n trees uses get_tree

#TOTALED DATA
#required for C5.0
gesture_tot_train$type<-as.factor(gesture_tot_train$type)

test_trees_rpart_t<- get_n_trees(npredictions,gesture_tot_train,nsamples,features_t,targetft_t,type=0)
test_trees_c50_t<- get_n_trees(npredictions,gesture_tot_train,nsamples,features_t,targetft_t,type=1)


#AVERAGED DATA '
#required for C5.0
avg_gesture_train$type<-as.factor(avg_gesture_train$type)

test_trees_rpart_a<- get_n_trees(npredictions,avg_gesture_train,nsamples,features_a,targetft_a,type=0)
test_trees_c50_a<- get_n_trees(npredictions,avg_gesture_train,nsamples,features_a,targetft_a,type=1)

```

```{r VISUZA VISUALIZING RPART TREES}
fancyRpartPlot(test_trees_rpart[[1]]$finalModel,sub="Tree #3",main = "RPART TREE")
```

```{r BINDS DECISION TREE PREDICTIONS INTO TABLE}
#SPREAD DATA
prediction_list_rpart <- bind_predictions(npredictions,test=head(gest_combine_test_no_type,nsamples),test_trees_rpart)

prediction_list_c50 <- bind_predictions(npredictions,test=head(gest_combine_test_no_type,nsamples),test_trees_c50)

#TOTALED DATA
#returns multiple decision tree predictions in one data frame
prediction_list_rpart_t <- bind_predictions(npredictions,test=head(gesture_tot_test_notype,nsamples),test_trees_rpart_t)
prediction_list_c50_t <- bind_predictions(npredictions,test=head(gesture_tot_test_notype,nsamples),test_trees_c50_t)

#average DATA
prediction_list_rpart_a <- bind_predictions(npredictions,test=head(avg_gesture_test_no_type,nsamples),test_trees_rpart_a)
prediction_list_c50_a <- bind_predictions(npredictions,test=head(avg_gesture_test_no_type,nsamples),test_trees_c50_a)

```

```{r CREATE 1 TABLE OF MEAN VALUES}
#SPREAD DATA
means_tbl_rpart<-get_mean(prediction_list_rpart)
means_tbl_c50<-get_mean(prediction_list_c50)

#Totaled DATA
means_tbl_rpart_t<-get_mean(prediction_list_rpart_t)
means_tbl_c50_t<-get_mean(prediction_list_c50_t)

#AVG DATA
means_tbl_rpart_a<-get_mean(prediction_list_rpart_a)
means_tbl_c50_a<-get_mean(prediction_list_c50_a)
```


```{r ERROR RATE CALCULATION}
#spread data
print("SPREAD DATA:")
error.rate_rpart = round(mean(head(gest_combine_test$type,nsamples) !=  means_tbl_rpart$means),2)
error.rate_c50 = round(mean(head(gest_combine_test$type,nsamples) !=  means_tbl_c50$means),2)

print("SPREAD ERROR RATE: ")
error.rate_rpart
error.rate_c50


#Totaled data
error.rate_rpart_t = round(mean(head(gesture_tot_test$type,nsamples) !=  means_tbl_rpart$means),2)
error.rate_c50_t = round(mean(head(gesture_tot_test$type,nsamples) !=  means_tbl_c50$means),2)


print("TOTALED ERROR RATE: ")
error.rate_rpart_t
error.rate_c50_t
print(" ")

#average data
error.rate_rpart_a = round(mean(head(avg_gesture_test$type,nsamples) !=  means_tbl_rpart_a$means),2)
error.rate_c50_a = round(mean(head(avg_gesture_test$type,nsamples) !=  means_tbl_c50_a$means),2)


print("AVG ERROR RATE: ")
error.rate_rpart_a
error.rate_c50_a
```

